### 2022-08-05 figures/fig-evolution-fst.tex

% !TEX program=lualatex
\RequirePackage{luatex85}
\documentclass[tikz,border=10,10pt]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{unicode-math}
\usepackage{fontspec}
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage{xcolor}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning}
\tikzset{
  font={\fontsize{9pt}{11}\selectfont}
}

\setmainfont{TeX Gyre Heros}
\setmathfont{Latin Modern Math}

\definecolor{colorR}{RGB}{228,26,28}    % RED
\definecolor{colorB}{RGB}{55,126,184}   % BLUE
\definecolor{colorG}{RGB}{77,175,74}    % GREEN
\definecolor{colorP}{RGB}{152,78,163}   % PURPLE
\definecolor{colorO}{RGB}{255,127,0}    % ORANGE
\definecolor{colorY}{RGB}{255,255,51}   % YELLOW
\definecolor{colorBn}{RGB}{166,86,40}   % BROWN
\definecolor{colorPk}{RGB}{247,129,191} % PINK
\definecolor{colorGy}{RGB}{153,153,153} % GRAY

\tikzstyle{line}=[draw, -stealth', very thick]
\tikzstyle{block}=[circle,fill=colorB!50,on grid]
\tikzstyle{lab}=[]
\tikzstyle{w}=[lab,midway]
\tikzstyle{e}=[lab,midway,auto=false,fill=colorP!50,font=\scriptsize]

\begin{document}
% \begin{tikzpicture}[node distance=25mm, auto]
\begin{tikzpicture}[node distance=25mm, auto,
	dottedline/.style = {ultra thick, loosely dotted,shorten >=1mm, shorten <=1mm}]
%%% MG94 marginal substitution FST
\node (titlea) {\textbf{a) Substitution}};

\node[block,below=2cm of titlea,fill=colorG!50] (start_sub) {start};
\node[block,right=15mm of start_sub] (S_sub) {S};
\node[block,right=60mm of S_sub] (M_sub) {M};
\node[block,fill=colorR!50,right=15mm of M_sub] (end_sub) {end};

\draw[line] (start_sub) -- (S_sub);
\draw[line] (S_sub) to[bend right=22.5] (M_sub);
% \draw[line] (S_sub) -- (M_sub);
\draw[line] (M_sub) -- (end_sub);

% \draw[line] (M_sub) to[bend right=85] node[e,pos=0.5] {X\textrightarrow{}Y} (S_sub);
% \draw[line] (M_sub) to[bend right=55] node[e,pos=0.5] {AAA1\textrightarrow{}A} (S_sub);
\draw[line] (M_sub) to[bend right=28.5] (S_sub);
\draw[line] (M_sub) to[bend right=16.5] (S_sub);
\draw[line] (M_sub) to[bend right=25.5] (S_sub);
\draw[line] (M_sub) to[bend right=19.5] (S_sub);
\draw[line] (M_sub) to[bend right=22.5] node[e,pos=0.5] {XXX\textrightarrow{}YYY} (S_sub);
% \draw[line] (M_sub) to[bend right=22.5] node[e,pos=0.5] {AAA1\textrightarrow{}C} (S_sub);
% \draw[line] (M_sub) to[bend right=8] node[e,pos=0.5] {AAA1\textrightarrow{}G} (S_sub);

% \draw[line] (M_sub) to[bend right=-8] node[e,pos=0.5] {AAA1\textrightarrow{}T} (S_sub);
% \draw[line] (M_sub) to[bend right=-22.5] node[e,pos=0.5] {AAA2\textrightarrow{}A} (S_sub);
% \draw[line] (M_sub) to[bend right=-65] node[e,pos=0.5] {TTT3\textrightarrow{}T} (S_sub);

% \draw[dottedline] (5,-3) -- node[below] {} + (0, -9mm);

%%% Legend

\node[below=2cm of start_sub,text width=50mm,anchor=north west] (sequences) {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Sequences}}\\
    \hline
	X & input nucleotides\\
	Y & intermediate nucleotides\\
	Z & output nucleotides\\
	\O & nothing/empty sequence\\
\end{tabular}
};

\node[above right=-1.4cm and 5mm of sequences,text width=50mm] {
\begin{tabular}{r@{\,: }l}
\multicolumn{2}{l}{\textbf{Parameters}}\\
\hline
	$g$ & gap open weight\\
	$e$ & gap extension weight\\
\end{tabular}
};

%%% Indel FST
\node[right=10cm of titlea] (titleb) {\textbf{b) Insertion-Deletion}};

\node[block,right=12cm of start_sub,fill=colorG!50] (start1) {start};
\node[block,below=15mm of start1] (S1) {S};
\node[block,below=of S1] (I1) {I};
\node[block,right=of I1] (U1) {U};
\node[block,above=of U1] (W1) {W};
\node[block,above=20mm of W1] (D1) {D};
\node[block,right=of D1] (V1) {V};
\node[block,right=of W1] (M1) {M};
\node[block,fill=colorR!50,right=15mm of M1] (end1) {end};

\draw (D1) +(0mm,-20mm) coordinate (C1);
\draw (D1) +(50mm,-10mm) coordinate (C2);

\draw[line] (start1) -- (S1);
\draw[line] (S1) -- node[w] {$g$} (I1);
\draw[line] (S1) -- node[w] {$1-g$} (W1);
\draw[line] (U1) to[out=150,in=30]  node[w,above] {$e$} (I1);
\draw[line] (U1) -- node[w,pos=0.1,right] {$1-e$} (W1);
\draw[line] (W1) -- node[w] {$g$} (D1);
\draw[line] (W1) -- node[w] {$1-g$} (M1);
\draw[line] (V1) to[out=150,in=30] node[w,above] {$e$} (D1);
\draw[line] (D1) to[out=60,in=90] (end1 |- D1.south) to (end1);
\draw[line] (V1) -- node[w,swap] {$1-e$} (M1);
\draw[line] (M1) --  (end1);

\draw[line] (I1) to[bend right=20] (U1);
\draw[line] (I1) to[bend right=25] (U1);
\draw[line] (I1) to[bend right=35] (U1);
\draw[line] (I1) to[bend right=40] (U1);
\draw[line] (I1) to[bend right=30] node[e] {\O\textrightarrow{}Z} (U1);

\draw[line] (D1) to[bend right=20] (V1);
\draw[line] (D1) to[bend right=25] (V1);
\draw[line] (D1) to[bend right=35] (V1);
\draw[line] (D1) to[bend right=40] (V1);
\draw[line] (D1) to[bend right=30] node[e] {Y\textrightarrow{}\O} (V1);

\draw[line] (M1) to[bend left=39] (S1);
\draw[line] (M1) to[bend left=42] (S1);
\draw[line] (M1) to[bend left=48] (S1);
\draw[line] (M1) to[bend left=51] (S1);

\draw[line] (M1) to[bend left=45] node[e,pos=0.25] {YYY\textrightarrow{}ZZZ} (S1);

% \node[below left=7.5mm and 1mm of I1.east,text width=50mm,anchor=north west] {
% \begin{tabular}{r@{\,: }l}
% \multicolumn{2}{l}{}\\
% %\hline
% 	X & input nucleotides\\
% 	Y & intermediate nucleotides\\
% 	Z & output nucleotides\\
% 	\O & nothing/empty sequence\\
%     \hline
% 	$g$ & gap open weight\\
% 	$e$ & gap extension weight\\
% \end{tabular}
% };

\end{tikzpicture}
\end{document}