### 2022-08-05 pairwise_appnote.tex

\include{header.tex}

% document begins here
\begin{document}
% \vspace*{0.35in}

% title goes here:
\begin{flushleft}
{\Large\textbf{COATi: }}
% codon-aware pairwise aligner + artifacts
\newline
% authors go here:
\\
Juan J. Garcia Mesa\textsuperscript{1,2},
Reed A. Cartwright\textsuperscript{1,3}
\\
\bigskip
1 Biodesign Institute, Arizona State University
\\
2 Ira A. Schools of Engineering, Arizona State University
\\
3 School of Life Sciences, Arizona State University
\\
\bigskip
* corresponding@author.mail

\end{flushleft}

% \abstract{Abstract}
\begin{abstract}
\noindent \textbf{Summary:} COATi is a statistical codon-aware pairwise aligner
that supports complex insertion-deletion models and is able to handle artifacts
present in genomic data.\\  % abstract summary can be longer.
\textbf{Availability:} The source code, along with documentation, for COATi is
freely available on GitHub: \url{https://github.com/CartwrightLab/coati} and is
implemented in C++.\\
\textbf{Supplementary information:} %\green{TODO}.
\end{abstract}

% now start line numbers
\linenumbers

\input{introduction}

\input{description2}

\input{results}

\section{Conclusions/Discussion}

COATi is an FST-based application that can find the optimal alignment between a
pair of sequences in the presence of artifacts using a statistical model.
% \green{(continue)}

\section*{Acknowledgments}

The author(s) would like to thank \green{person1}, \green{person2}, and
\green{number} reviewers for their helpful suggestions.

\subsection*{Funding}
This research was founded by an NSF-IIBR grant (\green{grant number}).\\

\noindent \textit{Conflict of interest:} none declared.

% \section{References}
% \printbibliography

\nolinenumbers

\end{document}