### 2022-08-05 results.tex

\section{Results}
% prelim data comparison?

Using 4000 human genes and their gorilla homologous pairs from the ENSEMBL
database \parencite{ensembl_hubbard_2002} we simulated a data set of pairwise
alignments with empirical gap patterns.
We used the data set to evaluate the accuracy of popular cutting edge aligners
Clustal$\Omega$ v1.2.4 \parencite{clustal_omega_sievers_2011}, MACSE v2.06
\parencite{ranwez_macse_2011}, MAFFT v7.407 \parencite{mafft_katoh_2002}, and
PRANK v.170427 \parencite{prank_loytynoja_2014} together with COATi.

After downloading, 1660 alignments contained gaps identified by at least one
aligner.
We randomly introduced gap patterns extracted from all five methods into the
2340 initially ungapped sequence pairs to generate the true alignments.
Alignment accuracy was measured using the distance metric $d_{seq}$
\parencite{metrics_blackburne_whelan_2011} between simulated and inferred
alignments.
In addition, accuracy of positive and negative selection was calculated
using the $F_1$ score by estimating $k_s$ and $k_a$ statistics.
% by estimating $k_s$ and $k_a$ statistics.

% % Software comparison table
% \begin{table}[!ht]
% \centering
%     %\frame{
%     \input{figures/table-comp}%}
% 	\caption{Accuracy of COATi, PRANK, MAFFT, Clustal$\Omega$, and MACSE,
%             on 2340 simulated sequence pairs. Perfect alignments have
%             ($d_{seq}=0$), best alignments have lowest $d_{seq}$, and imperfect
%             alignments have $d_{seq}>0$ when at least one aligner found a
%             perfect alignment. Best values are highlighted in blue.}
% 	\label{table:comp}
% \end{table}

COATi was significantly more accurate (lower $d_{seq}$) than other aligners; all
p-values were equal (MAFFT) or less than $1.714$x$10^{-8}$ according to the
one-tailed Wilcoxon signed rank test.
In addition, COATi produced more perfect alignments, less imperfect alignments,
and had a higher positive (97.3\%) selection accuracy.
MACSE (allows frameshifts) and MAFFT (using DNA) performed similarly with an
average alignment error an order of magnitude larger than COATi and the accuracy
retrieving positive selection of 81.5\% and 85.8\% respectively.
PRANK (using codons) and Clustal$\Omega$ (using amino acid translations) also
obtained similar results with an average alignment error two orders of magnitude
larger than COATi and a 87.3\% and 69.1\% accuracy retrieving positive selection.
The accuracy retrieving negative selection was similar across all five
aligners.

% (Table \ref{table:comp}).