### 2022-08-08 description2.tex

\section{Description}

% Summarize pair-HMMs. Brief, comment typically used but move onto FST right away
Statistical alignment is typically performed using pairwise hidden Markov
models (pair-HMMs).
% This computational machines consist of two output tapes and a set of states that
% emit symbols onto one or both tapes.
Pair-HMMs have the ability to rigorously model molecular sequence evolution and
can calculate the probability that two sequences are related, represented
$P(X, Y)$ \parencite{yoon_2009_hmm}.
However, a limitation of pair-HMMs is the ability to only model evolution of two
related sequences from an unknown ancestor.
Finite-state transducers (FSTs) have the ability to calculate the probability
that a sequence $Y$ evolved from sequence $X$, represented $P(Y | X)$.
FSTs share similar computational benefits as pair-HMMs in addition to well
established algorithms for combining them in different ways
\parencite{bradley2007transducers}.
A powerful and versatile algorithm is composition, which consists of sending the
output one FST as the input of a second FST.
The FST model implemented in COATi is designed by composing smaller FSTs, each
representing a specific process.

% Evolution FST
Pairwise alignment in COATi is implemented via the Evolution FST (Fig.
\ref{fig:evolution-fst}), based on existing transducers (e.g.
\cite{holmes2001evolutionary}).
The Evolution FST is formed by composing a substitution FST that encodes a 64x64
codon model (Fig. \ref{fig:evolution-fst}-a) and an indel FST that models
insertions and deletions, including frameshifts (Fig. \ref{fig:evolution-fst}-b).
The innovation of the Evolution FST with respect to other transducers is the
combination of a codon substitution model that allows stop codons with gaps that
can occur at any position of any length.

\begin{figure}[h!]
\begin{framed}
\centering
    \includegraphics[width=\textwidth]{fig-evolution-fst.pdf}
    \caption{The Evolution FST is assembled by composing a substitution FST and
    an indel FST. Each node represents a state in an FST while arcs display
    possible transitions between states (and their weights). Absorption and
    emission of symbols occurs between states. (a) The substitution FST
    encodes a 64x64 codon substitution model with 64 arcs from M to S. (b)
    The indel FST allows for insertions (I to U) and deletions (D to V).
    Contiguous insertions and deletions are always arranged for insertions to
    precede deletions to limit equivalent alignments.}
    \label{fig:evolution-fst}
\end{framed}
\end{figure}

% FST implementation
% A path through the FST that \green{results} from composing both sequences with
% the Evolution FST represent a pairwise alignment.
% To find the most \green{probable/likely} alignment we must find the shortest
% path through.
The alignment FST is the result from composing both sequences with the Evolution
FST.
Any path through the alignment FST represents a pairwise alignment, while the
shortest path corresponds to the best alignment.
All FST operations including creation of models, composition, search for the
shortest path, and other optimization algorithms are performed using the C++
openFST library \parencite{allauzen2007openfst}.

% However, composing large FSTs is an expensive operation and can be prohibitive.
% Despite the existence of efficient C++ FST libraries (e.g. openFST
% \cite{allauzen2007openfst}), runtime is still limiting for sequence pairs longer
% than a few thousand nucleotides each.
% To solve this issue, the search for an optimal path (alignment) is
% reformulated as a dynamic programming problem.

% Evolution FST + Q matrix
% Marginal model? - Probably not
% DP - probably not
% 1 figure: FST model? - results table?