### 2022-11-15 suppl.tex

\include{header.tex}

\begin{document}

{\large\textbf{Supplementary material}}

\begin{table}[!ht]
\centering
 \frame{\input{figures/table-comp}}
 \caption{Accuracy of COATi fst, COATi marginal, PRANK, MAFFT, Clustal$\Omega$,
         and MACSE, on 5399 simulated sequence pairs. Perfect alignments have
         ($d_{seq}=0$), best alignments have lowest $d_{seq}$, and imperfect
         alignments have $d_{seq}>0$ when at least one aligner found a
         perfect alignment.}
 \label{table:comp}
\end{table}

\end{document}