### 2023-05-04 supplementary_materials.Rmd

---
title: "Supplementary Materials"
#author: "Juan J. García Mesa"
#date: "`r Sys.Date()`"
geometry: margin=1.8cm
output:
    pdf_document:
        fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
library(knitr)

summary_table = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # extract dseq values for each aligner
    coati = results$dseq[which(results$aligner == coati_model)]
    clustalomega = results$dseq[which(results$aligner == "clustalo")]
    macse = results$dseq[which(results$aligner == "macse")]
    mafft = results$dseq[which(results$aligner == "mafft")]
    prank = results$dseq[which(results$aligner == "prank")]

    dseq_summary = data.frame(coati = mean(coati),
                      prank = mean(prank),
                      mafft = mean(mafft),
                      clustalo = mean(clustalomega),
                      macse = mean(macse), row.names = "$d_{seq}$")

    ############################################################################
    # one-side wilcoxon rank test
    wtp = wilcox.test(coati, prank)
    wtmf = wilcox.test(coati, mafft)
    wtc = wilcox.test(coati, clustalomega)
    wtmc = wilcox.test(coati, macse)
    wilcoxon = data.frame(coati = NA, prank = wtp$p.value, mafft = wtmf$p.value,
                          clustalo = wtc$p.value, macse = wtmc$p.value,
                          row.names = "Wilcoxon rank test")
    stats_table = rbind(dseq_summary, wilcoxon)

    ############################################################################
    # perfect, best, and imperfect alignments
    source("supplementary_materials/scripts/number_alignments.R")
    alns = num_alns(results_file)

    # fix row and column names and add to stats table
    row.names(alns)[which(row.names(alns) == coati_model)] = "coati"
    colnames(alns) = c("Perfect alignments", "Best alignments", "Imperfect alignments")
    stats_table = rbind(stats_table, t(alns))

    ############################################################################
    # selection (ka ks)
    source("supplementary_materials/scripts/kaks.R")

    aligners = c(coati_model, "prank", "mafft", "clustalo", "macse")
    selection = data.frame("F1.pos" = double(),
                           "F1.neg" = double())

    # calculate F1 score for positive and negative selection for each aligner
    for(model in aligners) {
        model_rows = results[which(results$aligner == model), ]
        pos = ps_accuracy(model_rows$ref_omega, model_rows$aln_omega)
        neg = ns_accuracy(model_rows$ref_omega, model_rows$aln_omega)

        selection[nrow(selection) + 1, ] = c(pos, neg)
    }

    # fix row and column names and add to stats table
    colnames(selection) = c("F1-score pos selection", "F1-score neg selection")
    rownames(selection) = c("coati", "prank", "mafft", "clustalo", "macse")
    stats_table = rbind(stats_table, t(selection))
    colnames(stats_table) = c(coati_model, "PRANK", "MAFFT", "ClustalOmega", "MACSE")

    # display stats table
    stats_table
}

```

```{r echo = FALSE, results = 'asis'}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")

results = "supplementary_data/tri-mg/results/results_summary.csv"
print(kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7719 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/tri-ecm/results/results_summary.csv"
print(kable(summary_table(results, "tri-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7678 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/mar-mg/results/results_summary.csv"
print(kable(summary_table(results, "mar-mg"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7661 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))

results = "supplementary_data/mar-ecm/results/results_summary.csv"
print(kable(summary_table(results, "mar-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7710 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq}$>0 when at least one aligner found a perfect alignment.")))
```

\newpage

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = "\\label{fig:dseq-tri-mg} Comparison of $d_{seq}$ values between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/tri-mg/results/results_summary.csv", "tri-mg")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-tri-ecm} Comparison of $d_{seq}$ values between COATi codon-triplet-ecm and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/tri-ecm/results/results_summary.csv", "tri-ecm")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-mar-mg} Comparison of $d_{seq}$ values between COATi codon-marginal-mg and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/mar-mg/results/results_summary.csv", "mar-mg")
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap =  "\\label{fig:dseq-mar-ecm} Comparison of $d_{seq}$ values between COATi codon-marginal-ecm and PRANK, MAFFT, ClustalOmega, and MACSE."}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
plot_dseq_main("supplementary_data/mar-ecm/results/results_summary.csv", "mar-ecm")
```

\newpage

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.cap="\\label{fig:figs} Base calling error FST. Arcs from M to S generate matches; however, here they can introduce single-nucleotide errors, which can generate stop codon artifacts."}
knitr::include_graphics(path = "figures/fig-base-calling-error.pdf")
```