### 2023-05-12 supplementary_materials.Rmd
Update comp table and supplementary materials

---
title: "Supplementary Materials"
#author: "Juan J. García Mesa"
#date: "`r Sys.Date()`"
geometry: margin=1.8cm
output:
    pdf_document:
        fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
library(knitr)

summary_table = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # extract dseq values for each aligner
    coati = results$dseq[which(results$aligner == coati_model)]
    clustalomega = results$dseq[which(results$aligner == "clustalo")]
    macse = results$dseq[which(results$aligner == "macse")]
    mafft = results$dseq[which(results$aligner == "mafft")]
    prank = results$dseq[which(results$aligner == "prank")]

    dseq_summary = data.frame(coati = mean(coati, na.rm = TRUE),
                      prank = mean(prank, na.rm = TRUE),
                      mafft = mean(mafft, na.rm = TRUE),
                      clustalo = mean(clustalomega, na.rm = TRUE),
                      macse = mean(macse, na.rm = TRUE),
                      row.names = "$d_{seq}$")

    ############################################################################
    # perfect, best, and imperfect alignments
    source("supplementary_materials/scripts/number_alignments.R")
    alns = num_alns(results_file)

    # fix row and column names and add to stats table
    row.names(alns)[which(row.names(alns) == coati_model)] = "coati"
    colnames(alns) = c("Perfect alignments", "Best alignments", "Imperfect alignments")
    stats_table = rbind(dseq_summary, t(alns))

    ############################################################################
    # selection (ka ks)
    source("supplementary_materials/scripts/kaks.R")

    aligners = c(coati_model, "prank", "mafft", "clustalo", "macse")
    selection = data.frame("F1.pos" = double(),
                           "F1.neg" = double())

    # calculate F1 score for positive and negative selection for each aligner
    for(model in aligners) {
        model_rows = results[which(results$aligner == model), ]
        pos = ps_accuracy(model_rows$ref_omega, model_rows$aln_omega)
        neg = ns_accuracy(model_rows$ref_omega, model_rows$aln_omega)

        selection[nrow(selection) + 1, ] = c(pos, neg)
    }

    # fix row and column names and add to stats table
    colnames(selection) = c("F1-score pos selection", "F1-score neg selection")
    rownames(selection) = c("coati", "prank", "mafft", "clustalo", "macse")
    stats_table = rbind(stats_table, t(selection))
    colnames(stats_table) = c(coati_model, "PRANK*", "MAFFT", "ClustalOmega", "MACSE")
    stats_table = stats_table[, c(coati_model, "MAFFT", "PRANK*", "MACSE", "ClustalOmega")]

    # display stats table
    stats_table
}

wilcoxon_test = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # extract dseq values for each aligner
    coati = results$dseq[which(results$aligner == coati_model)]
    clustalomega = results$dseq[which(results$aligner == "clustalo")]
    macse = results$dseq[which(results$aligner == "macse")]
    mafft = results$dseq[which(results$aligner == "mafft")]
    prank = results$dseq[which(results$aligner == "prank")]

    # one-side paired wilcoxon rank test
    wtp = wilcox.test(x = coati, y = prank, paired = TRUE, alternative = "less")
    wtmf = wilcox.test(x = coati, y = mafft, paired = TRUE, alternative = "less")
    wtc = wilcox.test(x = coati, y = clustalomega, paired = TRUE, alternative = "less")
    wtmc = wilcox.test(x = coati, y = macse, paired = TRUE, alternative = "less")
    w = c(wtp$p.value, wtmf$p.value, wtc$p.value, wtmc$p.value)
    w
}
```

```{r echo = FALSE, results = 'asis'}
library(kableExtra)
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
results = "supplementary_data/tri-mg/results/results_summary.csv"
k = kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7719 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment. "), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 50 empty alignments, calculations are based on 7669 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/tri-ecm/results/results_summary.csv"
k = kable(summary_table(results, "tri-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7678 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 41 empty alignments, calculations are based on 7637 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/mar-mg/results/results_summary.csv"
k = kable(summary_table(results, "mar-mg"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7661 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment. "), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 50 empty alignments, calculations are based on 7611 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
#
results = "supplementary_data/mar-ecm/results/results_summary.csv"
k = kable(summary_table(results, "mar-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7710 simulated sequence pairs. Perfect alignments have $d_{seq}$ = 0, best alignments have lowest $d_{seq}$, and imperfect alignments have $d_{seq} > 0$ when at least one aligner found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 35 empty alignments, calculations are based on 7675 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```

\newpage

```{r echo = FALSE}
results = "supplementary_data/tri-mg/results/results_summary.csv"
pval_tri_mg = max(wilcoxon_test(results, "tri-mg"))
results = "supplementary_data/tri-ecm/results/results_summary.csv"
pval_tri_ecm = max(wilcoxon_test(results, "tri-ecm"))
results = "supplementary_data/mar-mg/results/results_summary.csv"
pval_mar_mg = max(wilcoxon_test(results, "mar-mg"))
results = "supplementary_data/mar-ecm/results/results_summary.csv"
pval_mar_ecm = max(wilcoxon_test(results, "mar-ecm"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-mg} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-mg/plot_distance.csv", "tri-mg"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-ecm} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-ecm/plot_distance.csv", "tri-ecm"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-mg} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-mg/plot_distance.csv", "mar-mg"))
```

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-ecm} Comparison of log-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-ecm/plot_distance.csv", "mar-ecm"))
```

\newpage

```{r echo = FALSE, fig.width = 6, fig.height = 6, fig.cap="\\label{fig:figs} Base calling error FST. Arcs from M to S generate matches; however, here they can introduce single-nucleotide errors, which can generate stop codon artifacts."}
knitr::include_graphics(path = "figures/fig-base-calling-error.pdf")
```