### 2023-12-21 figures/fig-subst-fst.tex

\definecolor{colorR}{RGB}{228,26,28}    % RED
\definecolor{colorB}{RGB}{55,126,184}   % BLUE
\definecolor{colorG}{RGB}{77,175,74}    % GREEN
\definecolor{colorP}{RGB}{152,78,163}   % PURPLE
\definecolor{colorO}{RGB}{255,127,0}    % ORANGE
\definecolor{colorY}{RGB}{255,255,51}   % YELLOW
\definecolor{colorBn}{RGB}{166,86,40}   % BROWN
\definecolor{colorPk}{RGB}{247,129,191} % PINK
\definecolor{colorGy}{RGB}{153,153,153} % GRAY

\tikzstyle{line}=[draw, -stealth', very thick]
\tikzstyle{block}=[circle,fill=colorB!50,on grid]
\tikzstyle{lab}=[]
\tikzstyle{w}=[lab,midway]
\tikzstyle{e}=[lab,midway,above=0.2mm,auto=false,font=\scriptsize]

% \begin{document}
% \begin{tikzpicture}[node distance=25mm, auto]
\begin{tikzpicture}[node distance=25mm, auto,
	dottedline/.style = {ultra thick, loosely dotted,shorten >=1mm, shorten <=1mm}]
%%% MG94 marginal substitution FST
% \node (titlea) {};

\node[block,fill=colorG!50] (start_sub) {start};
\node[block,right=15mm of start_sub] (S_sub) {M};
\node[block,right=75mm of S_sub] (M_sub) {S};
\node[block,fill=colorR!50,right=15mm of M_sub] (end_sub) {end};

% AAA->AAA
\node[block,above right=12mm and 20mm of S_sub,minimum size=15pt] (s2) {};
\node[block,above right=3mm and 20mm of s2,minimum size=15pt] (s1) {};
% % AAA->AAC
\node[block,above right=4mm and 20mm of S_sub,minimum size=15pt] (s4) {};
\node[block,below right=-1mm and 20mm of s4,minimum size=15pt] (s3) {};
% TTT->TTT
\node[block,below right=6mm and 20mm of S_sub,minimum size=15pt] (s122) {};
\node[block,below right=3mm and 20mm of s122,minimum size=15pt] (s121) {};

\draw[line] (start_sub) -- (S_sub);
\draw[line] (S_sub) to[bend right=40] (M_sub);
\draw[line] (M_sub) -- (end_sub);

% AAA->AAA
% \draw[line] (M_sub) to[bend right=15] node[w,pos=0.7,above right=3mm and 1mm,font=\scriptsize,rotate=-20] {$P(AAA|AAA)$} (s1);
\draw[line] (M_sub) to[bend right=15] node[e,pos=0.4,rotate=-20] {A:A/$P(AAA|AAA)$} (s1);
\draw[line] (s1) to[bend right=15] node[e,pos=0.5] {A:A} (s2);
\draw[line] (s2) to[bend right=15] node[e,pos=0.5] {A:A} (S_sub);

% % AAA->AAC
\draw[line] (M_sub) to[bend right=5] node[e,pos=0.5,rotate=-8,below=-5mm] {A:A/$P(AAC|AAA)$} (s3);
\draw[line] (s3) to[bend right=8] node[e,pos=0.5] {A:A} (s4);
\draw[line] (s4) to[bend right=8] node[e,pos=0.5] {A:C} (S_sub);

\node[below=9mm of s1,font=\Huge] {$\vdots$};
\node[below=5mm of s2,font=\Huge] {$\vdots$};

% TTT->TTT
% \draw[line] (M_sub) to[bend right=-5] node[w,pos=0.5,above right=1mm and -11mm, rotate=5,font=\scriptsize] {$P(TTT|TTT)$} (s121);
\draw[line] (M_sub) to[bend right=-5] node[e,pos=0.5,rotate=9] {T:T/$P(TTT|TTT)$} (s121);
\draw[line] (s121) to[bend right=-5] node[e,pos=0.5] {T:T} (s122);
\draw[line] (s122) to[bend right=-5] node[e,pos=0.4] {T:T} (S_sub);
\end{tikzpicture}