### 2023-12-23 figures/fig-fst-base-calling-new.tex
Add new COATi FST figure.

% !TEX program=lualatex
\documentclass[tikz]{standalone}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[warnings-off={mathtools-colon,mathtools-overbracket}]{unicode-math}
\usepackage{fontspec}
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage{xcolor}

\usetikzlibrary{shapes,arrows.meta,positioning,calc}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.text}
\usetikzlibrary{hobby,spath3,intersections}

\setmainfont{TeX Gyre Heros}
\setmathfont{Latin Modern Math}

% \tikzset{
%   font={\fontsize{6}{7}\selectfont}
% }

%%% Okabe Ito set
\definecolor{colorO}{RGB}{230,159,0}
\definecolor{colorSb}{RGB}{86,180,233}
\definecolor{colorBg}{RGB}{0,158,115}
\definecolor{colorY}{RGB}{240,228,66}
\definecolor{colorBu}{RGB}{0,114,178}
\definecolor{colorV}{RGB}{213,94,0}
\definecolor{colorRp}{RGB}{204,121,167}
\definecolor{colorBk}{RGB}{0,0,0}
\definecolor{colorGy}{RGB}{153,153,153} % GRAY

% (86 mm) or double (178 mm) column width

\tikzset{
  end/.style={fill=none, text=black, inner sep=1pt},
  block/.style={circle,on grid,fill=colorBu,text=white,minimum width=2em,node distance=4em},
  line/.style={draw, -Stealth, thick},
  lab/.style={text = black, inner sep=1pt},
  title/.style={text = black, font=\bfseries}
}

\begin{document}
\fontsize{8}{9}\selectfont
\begin{tikzpicture}
%\useasboundingbox (0,0) rectangle +(178mm,178mm);

\node[end, anchor=west] at (0,0) (start) {start};
\node[end, anchor=east] at (86mm,0) (end) {end};

\node[block,right = of start] (m) {M}; 
\node[block,left = of end] (s) {S};

\draw[line] (start) to (m);
\draw[line] (s) to (end);

\coordinate (mid) at ($(s.center)!0.5!(m.center)$);

\draw[line] (m) to (s);

% \path[spath/save=sm] (s.center) circle (1em+0.2pt)
%                   (m.center) circle (1em+0.2pt);

\foreach \y [count=\yi] in {
  {A:A/$(1-u)$}, {A:C/$(u/3)$}, {A:G/$(u/3)$}, {A:T/$(u/3)$},
  {C:A/$(u/3)$}, {C:C/$(1-u)$}, {C:G/$(u/3)$}, {C:T/$(u/3)$}
} {
    \coordinate (y) at ($(mid) + (0, 15.5em) + \yi*(0, -1.75em)$);
    % \path[spath/save=arc] (s.center) to[curve through={(y)}] (m.center);

    % \tikzset{
    %   spath/split at intersections with={arc}{sm},
    %   spath/remove components={arc}{1,3}
    % }
    % \path[line,spath/use=arc];

    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=south] at (y) {\y};
}

\foreach \y [count=\yi] in {
  {G:A/$(u/3)$}, {G:C/$(u/3)$}, {G:G/$(1-u)$}, {G:T/$(u/3)$},
  {T:A/$(u/3)$}, {T:C/$(u/3)$}, {T:G/$(u/3)$}, {T:T/$(1-u)$}
} {
    \coordinate (y) at ($(mid) + (0, 0.25em) + \yi*(0, -1.75em)$);
    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=north] at (y) {\y};
}

\node[lab,fill=white] at (mid) {ε:ε/$1$};

\draw (start.west) + (0, 15.5em) node[title,anchor=west] (subfig-a) {a) Error FST};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\node[end, anchor=west] at (0,-70mm) (start) {start};
\node[end, anchor=east] at (86mm,-70mm) (end) {end};

\node[block,right = of start] (m) {M}; 
\node[block,left = of end] (s) {S};

\draw[line] (start) to (m);
\draw[line] (s) to (end);

\coordinate (mid) at ($(s.center)!0.5!(m.center)$);

\draw[line] (m) to (s);

\foreach \y [count=\yi] in {
  {A:A/$1$}, {C:C/$1$}, {G:G/$1$}, {T:T/$1$}
} {
    \coordinate (y) at ($(mid) + (0, 8.5em) + \yi*(0, -1.75em)$);
    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=south] at (y) {\y};
}

\foreach \y [count=\yi] in {
  {A:N/$1$}, {C:N/$1$}, {G:N/$1$}, {T:N/$1$}
} {
    \coordinate (y) at ($(mid) + (0, 0.25em) + \yi*(0, -1.75em)$);
    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=north] at (y) {\y};
}

\node[lab,fill=white] at (mid) {ε:ε/$1$};

\draw (start.west) + (0, 8.5em) node[title,anchor=west] (subfig-b) {b) Ambiguity FST};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\node[end, anchor=west] at (92mm,-3.5em) (start) {start};
\node[end, anchor=east] at (178mm,-3.5em) (end) {end};

\node[block,right = of start] (m) {M}; 
\node[block,left = of end] (s) {S};

\draw[line] (start) to (m);
\draw[line] (s) to (end);

\coordinate (mid) at ($(s.center)!0.5!(m.center)$);

\draw[line] (m) to (s);

\foreach \y [count=\yi] in {
  {A:N/$1$}, {A:A/$(1-u)$}, {A:C/$(u/3)$}, {A:G/$(u/3)$}, {A:T/$(u/3)$},
  {C:N/$1$}, {C:A/$(u/3)$}, {C:C/$(1-u)$}, {C:G/$(u/3)$}, {C:T/$(u/3)$}
} {
    \coordinate (y) at ($(mid) + (0, 19em) + \yi*(0, -1.75em)$);
    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=south] at (y) {\y};
}

\foreach \y [count=\yi] in {
  {G:N/$1$}, {G:A/$(u/3)$}, {G:C/$(u/3)$}, {G:G/$(1-u)$}, {G:T/$(u/3)$},
  {T:N/$1$}, {T:A/$(u/3)$}, {T:C/$(u/3)$}, {T:G/$(u/3)$}, {T:T/$(1-u)$}
} {
    \coordinate (y) at ($(mid) + (0, 0.25em) + \yi*(0, -1.75em)$);
    \path[line,shorten <=1em+0.2pt, shorten >=1em+0.3pt] (s.center) to[curve through={(y)}] (m.center);
    \node[lab,anchor=north] at (y) {\y};
}

\node[lab,fill=white] at (mid) {ε:ε/$1$};

\draw (start.west) + (0, 19em) node[title,anchor=west] (subfig-b) {c) Error $∘$ Ambiguity FST};

\end{tikzpicture}
\end{document}