### 2024-01-01 supplementary_materials.Rmd

---
title: "Supplementary Materials for \"COATi: statistical pairwise alignment of protein coding sequences\""
author: Juan J. Garcia Mesa, Ziqi Zhu, Reed A. Cartwright
output:
    pdf_document:
        fig_caption: yes
        keep_tex: true
bibliography: alignpair_letter.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE}
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tidyverse))

summary_table = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # extract dseq values for each aligner
    coati = results |> filter(aligner == coati_model) |> select(dseq)
    clustal = results |> filter(aligner == "clustalo") |> select(dseq)
    macse = results |> filter(aligner == "macse") |> select(dseq)
    mafft = results |> filter(aligner == "mafft") |> select(dseq)
    prank = results |> filter(aligner == "prank") |> select(dseq)

    stats_table = data.frame(dseq = c(mean(coati$dseq, na.rm = TRUE),
                          mean(mafft$dseq, na.rm = TRUE),
                          mean(prank$dseq, na.rm = TRUE),
                          mean(macse$dseq, na.rm = TRUE),
                          mean(clustal$dseq, na.rm = TRUE)), row.names = c(coati_model, "mafft", "prank", "macse", "clustalo"))

    ###########################################################################
    # perfect, best, and imperfect alignments
    source("supplementary_materials/scripts/number_alignments.R")
    alns = num_alns(results_file)

    # fix row and column names and add to stats table
    colnames(alns) = c("Perfect alns", "Best alns", "Imperfect alns")
    alns = alns[c(coati_model, "mafft", "prank", "macse", "clustalo"), ]
    stats_table = cbind(stats_table, alns)

    ############################################################################
    # selection (ka ks)
    source("supplementary_materials/scripts/kaks.R")

    aligners = c(coati_model, "mafft", "prank", "macse", "clustalo")
    selection = data.frame("F1.pos" = double(),
                           "F1.neg" = double())

    # calculate F1 score for positive and negative selection for each aligner
    for(model in aligners) {
        model_rows = results[which(results$aligner == model), ]
        pos = ps_accuracy(model_rows$ref_omega, model_rows$aln_omega)
        neg = ns_accuracy(model_rows$ref_omega, model_rows$aln_omega)

        selection[nrow(selection) + 1, ] = c(pos, neg)
    }

    # fix row and column names and add to stats table
    colnames(selection) = c("F1 pos selection", "F1 neg selection")
    rownames(selection) = c("coati", "mafft", "prank", "macse", "clustalo")
    stats_table = cbind(stats_table, selection)
    if(coati_model == "tri-mg") {
        coati_model = "Triplet-MG94"
    } else if(coati_model == "tri-ecm") {
        coati_model = "Triplet-ECM"
    } else if(coati_model == "mar-mg") {
        coati_model = "Marginal-MG94"
    } else if(coati_model == "mar-ecm") {
        coati_model = "Marginal-ECM"
    }
    rownames(stats_table) = c(coati_model, "MAFFT", "PRANK*", "MACSE", "ClustalOmega")

    # display stats table
    stats_table
}

wilcoxon_test = function(results_file, coati_model) {
    # read results_summary data
    results = read.csv(results_file)

    # extract dseq values for each aligner
    coati = results$dseq[which(results$aligner == coati_model)]
    clustalomega = results$dseq[which(results$aligner == "clustalo")]
    macse = results$dseq[which(results$aligner == "macse")]
    mafft = results$dseq[which(results$aligner == "mafft")]
    prank = results$dseq[which(results$aligner == "prank")]

    # one-side paired wilcoxon rank test
    wtp = wilcox.test(x = coati, y = prank, paired = TRUE, alternative = "less")
    wtmf = wilcox.test(x = coati, y = mafft, paired = TRUE, alternative = "less")
    wtc = wilcox.test(x = coati, y = clustalomega, paired = TRUE, alternative = "less")
    wtmc = wilcox.test(x = coati, y = macse, paired = TRUE, alternative = "less")
    w = c(wtp$p.value, wtmf$p.value, wtc$p.value, wtmc$p.value)
    w
}
```

```{r setup-table-results, echo = FALSE}
suppressMessages(library(kableExtra))
suppressMessages(library(knitr))
suppressMessages(library(tidyverse))
```

```{r table-results-1, echo = FALSE, results = 'asis'}
results = "supplementary_data/tri-mg/results/results_summary.csv"
k = kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7761 simulated sequence pairs. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 42 empty alignments, calculations are based on 7719 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```
```{r table-results-2, echo = FALSE, results = 'asis'}
results = "supplementary_data/tri-ecm/results/results_summary.csv"
k = kable(summary_table(results, "tri-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7761 simulated sequence pairs. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 69 empty alignments, calculations are based on 7692 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```
```{r table-results-3, echo = FALSE, results = 'asis'}
results = "supplementary_data/mar-mg/results/results_summary.csv"
k = kable(summary_table(results, "mar-mg"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7755 simulated sequence pairs. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 60 empty alignments, calculations are based on 7695 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```
```{r table-results-4, echo = FALSE, results = 'asis'}
results = "supplementary_data/mar-ecm/results/results_summary.csv"
k = kable(summary_table(results, "mar-ecm"), digits = 5, caption = paste("Accuracy of COATi codon-marginal-ecm, PRANK, MAFFT, ClustalOmega, and MACSE on 7767 simulated sequence pairs. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 49 empty alignments, calculations are based on 7718 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```
```{r table-results-5, echo = FALSE, results = 'asis'}
results = "supplementary_data/reverse/results/results_summary.csv"
k = kable(summary_table(results, "tri-mg"), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, and MACSE on 7798 simulated sequence pairs with gorilla as the reference. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(footnote(k, general = "PRANK produced 35 empty alignments, calculations are based on 7763 alignments.", footnote_as_chunk = TRUE, general_title = "*"))
```
```{r table-results-6, echo = FALSE, results = 'asis'}
source("supplementary_materials/scripts/results_gor_ref.R")
k = kable(gor_ref_results(), digits = 5, caption = paste("Accuracy of COATi codon-triplet-mg, PRANK, MAFFT, ClustalOmega, MACSE, and codon-triplet-mg with gorila as the reference on 4003 of the 7761 simulated sequence pairs where the gorilla sequence was simulated without early stop codons, incomplete codons, or ambiguous nucleotides. Perfect alignments have the same score as the true alignment, best alignments have lowest $d_{seq}$, and imperfect alignments have a different score than the true alignment when at least one method found a perfect alignment."), escape = FALSE) %>% kable_styling(latex_options = "HOLD_position")
print(k)
```

\newpage

```{r base-calling-error, echo = FALSE, fig.width = 6, fig.height = 6, fig.cap="\\label{fig:figs} Base calling error FST. Arcs from M to S generate matches; however, here they can introduce single-nucleotide errors, which can generate stop codon artifacts."}
knitr::include_graphics(path = "figures/fig-base-calling-error.pdf")
```

```{r aln-example, echo = FALSE, fig.show='hold', fig.align='center', fig.cap="\\label{fig:aln-example} FST of all possible alignments between `CTC' and `CTG' (top) and the best alignment, determined by the shortest path (bottom). Figures generated using the OpenFST library."}
knitr::include_graphics(path = c("figures/aln-example-graph.pdf", "figures/aln-example.pdf"))
```

```{r pvals, echo = FALSE}
results = "supplementary_data/tri-mg/results/results_summary.csv"
pval_tri_mg = max(wilcoxon_test(results, "tri-mg"))
results = "supplementary_data/tri-ecm/results/results_summary.csv"
pval_tri_ecm = max(wilcoxon_test(results, "tri-ecm"))
results = "supplementary_data/mar-mg/results/results_summary.csv"
pval_mar_mg = max(wilcoxon_test(results, "mar-mg"))
results = "supplementary_data/mar-ecm/results/results_summary.csv"
pval_mar_ecm = max(wilcoxon_test(results, "mar-ecm"))
results = "supplementary_data/reverse/results/results_summary.csv"
pval_reverse = max(wilcoxon_test(results, "tri-mg"))
```

```{r dseq1, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-mg} Comparison of log10-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-mg/plot_distance.csv", "tri-mg", save = TRUE))
knitr::include_graphics(path = "figures/dseq_plots_tri-mg.pdf")
```

```{r dseq2, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-tri-ecm} Comparison of log10-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_tri_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/tri-ecm/plot_distance.csv", "tri-ecm", save = TRUE))
knitr::include_graphics(path = "figures/dseq_plots_tri-ecm.pdf")
```

```{r dseq3, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-mg} Comparison of log10-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-mg and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_mg, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-mg/plot_distance.csv", "mar-mg", save = TRUE))
knitr::include_graphics(path = "figures/dseq_plots_mar-mg.pdf")
```

```{r dseq4, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-mar-ecm} Comparison of log10-transformed $d_{seq}$ data with pseudocounts between COATi codon-marginal-ecm and PRANK, MAFFT, ClustalOmega, and MACSE. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_mar_ecm, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/mar-ecm/plot_distance.csv", "mar-ecm", save = TRUE))
knitr::include_graphics(path = "figures/dseq_plots_mar-ecm.pdf")
```

```{r dseq5, echo = FALSE, fig.width = 6, fig.height = 6, fig.align = 'center', fig.cap = paste0("\\label{fig:dseq-reverse} Comparison of log10-transformed $d_{seq}$ data with pseudocounts between COATi codon-triplet-mg and PRANK, MAFFT, ClustalOmega, and MACSE with gorilla as the reference. COATi was significantly more accurate than other aligners; all p-values were $\\leq ", signif(pval_reverse, digits = 3), "$.")}
library(knitr)
source("supplementary_materials/scripts/plot_dseq.R")
suppressWarnings(plot_dseq_main("supplementary_data/reverse/plot_distance.csv", "reverse", save = TRUE))
knitr::include_graphics(path = "figures/dseq_plots_reverse.pdf")
```

\newpage

```{r k2p-dist, echo = FALSE, fig.cap="\\label{fig:k2p} K2P distances. Evolutionary distances between reference and retrieved alignments for each method calculated using the Kimura 2-parameter model. Notably, COATi performs best while the other tools tend to overestimate the evolutionary distance."}
knitr::include_graphics(path = "figures/k2p_distances.pdf")
```

<!-- \newpage -->

<!-- # Supplementary Methods -->

<!-- Ks and Ka represent the number of substitutions per synonymous and non-synonymous sites. The ratio of nonsynonymous (Ka) to synonymous (Ks) nucleotide substitution rates indicates the selective pressures acting on genes. If the ratio is significantly greater than 1, it suggests positive selective pressure, meaning that nonsynonymous substitutions occur more frequently than synonymous substitutions. A ratio around 1 can indicate either neutral evolution at the protein level or a mixture of positive and negative selective pressures. If the ratio is less than 1, it indicates a pressure to maintain protein sequence, known as purifying selection. -->
<!-- Ks and Ka are calculated using the R package seqinr v.4.2-30 [@seqinr]. -->

<!-- # References -->

<!-- <div id="refs"></div> -->